---
title: "Write Up - Christian Smith"
output: html_document
date: '2022-08-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Explanation

As I was working through this project and code I found myself doing quite a bit of cleaning. Whether it was me having to go back to include another variable, or adding in a new variable to account for time periods etc., I spent a fair amount of time wrangling the data. There was a lot of back tracking and rerunning the code in order to start to get it to a comfortable place where I felt the data was in a good spot for cleaning after working out many of the kinks. After splitting out the data by pre college scorecard release and post college scorecard release, I wanted to a basic quick view of the differences in median income and mean median income just as a way to get a sense of the changes between the two before I started to get into the meat of running models and and regressions. What surprised me was that it didn't show much of a change at all from pre and post in terms of median income 10 years after graduation. In fact it actually decreased from afterwards, which I would not have expected from the initial question and simply from intuition. So I started to think about reasons this might be in order to get variables for them or to be able to measure them. I thought about the differences in degrees and figured that it might also have a large part to do with median income not changing as the type of degree you get will dictate the industry you go into after graduation (for the most part, there are obviously some people who make a pivot). But then I thought about how the effects from that should balance out either way, and the question we want to look into is by school not my degree. So I thought about cost as well, knowing that the cost of the college also has an impact on whether people go to that school based on family wealth/in state opportunities etc. Obviously if Harvard or Yale were as cheap as in state schools then a lot more people would go, but again, the question we are trying to answer is whether the scorecard had an effect on people going to schools with higher median incomes, and being able to afford a certain school does not change. So when I went back to the data I figured there were a couple ways to measure it, we could look at the type of school as listed by the "CONTROL" variable, we could take a look at a certain split out between high and low earnings people and see if there are differences there based on control. We also could look at the interaction between these variables and see if those account for more explanation of the numbers. So as I started thinking about that, I played around with a lot of combinations based on the pre and post scorecard release, getting an idea of the relationship between certain variables in high and low income people, both before and after the scorecard. I also played with some controls and a couple interactions, but most of them weren't giving me a great r squared even with significant level p values. After going through all of these and not wanting to use them, I decided to take a look at a few basic graphs to see the trends over time, and what I saw was that Pre scorecard data remained fairly steady across time in regards to median income, but Post scorecard data showed a steady decline over time in terms of median income. Based on these plots below, I decided to do some comparisons of models Pre and Post scorecard now that I had a better sense of direction and where I should focus the analysis to go. Now I also did some analysis on those same variables but used the High and Low earning schools as well, but since the low earning schools had a pretty consistent median income over time, it wouldn't impact the trends we saw, so I just decided to use to whole Pre and Post sets.

```{r cars}
ggplot(data = Pre, aes(Day, median_earn)) +
  geom_smooth() + ggtitle("Pre Sept 2015")

ggplot(data = Post, aes(Day, median_earn)) +
  geom_smooth() + ggtitle("Post Sept 2015")
```

So after I got some better direction on how to handle the regression models I wanted to run, I began messing around with some models to see where the key relationships might be, and some again like before were significant, but the r squared was pretty low so I figured I would try a couple interactions and additions to see where I could possibly keep the significance level of the variables, but also get a better r squared, which as mentioned in the lectures shouldn't be super focused on, but I figured I would try and see what I could get and it ended up working out pretty good. The models below are the ones I focused mainly on. Of the three comparisons I looked at with etables, two of the three had much higher r squared which I was originally looking for. Those models were the two 'another' models comparison and the 't77' & 'g2' comparisons. Based on these two I looked at the wald results for them to see if there was any more insight there that I could grab for certain variables, which again it showed high p value/significance at a high confidence level on the varaibles I took a look at. Even though I did some playing around with the variables to see if any were indeed equal to zero, I only sw high values as well as a very significant p value for the tests which shows that all of these variables I am using are significant and should stay in the model. So after all this I decided that I felt the best model to include would be the one that has 'sdindex' as well as the interaction between type of school and high median earning schools. I felt that this interaction was worth including as it would give us a sense of how the type of school and weather it's a high earning school should be included.


```{r pressure, echo=FALSE}

t7 <- feols(median_earn ~ sdindex + CONTROL, data = Pre)
t77 <- feols(median_earn ~ sdindex + High, data = Pre)
g24  <- feols(median_earn ~ sdindex + CONTROL, data = Post)
g2 <- feols(median_earn ~ sdindex + High, data = Post)

another <- feols(median_earn ~ sdindex + CONTROL * High, data = Pre)
another2 <- feols(median_earn ~ sdindex + CONTROL * High, data = Post)

etable(another, another2)
```

so based on all this information, what this tells me is that before the scorecard was released, a one unit change in the standard deviation index we calculated will result in an increase of earnings by 35.19, a one unit increase on a high earning school type of school, (we will use 1 for public as the example) will increase earnings by 1,066.3, a one unit increase in the type of school so public, private non profit, and private for profit in that order will decrease earnings 814, and for those in high earning schools, they will increase their earnings by 15,270.7. And after the college score card released, a one unit change in the standard deviation index we calculated will result in an increase of earnings by 138.9, a one unit increase on a high earning school type of school, (we will use 1 for public as the example) will increase earnings by 1,007.2, a one unit increase in the type of school so public, private non profit, and private for profit in that order will decrease earnings by 791.3, and for those in high earning schools, they will increase their earnings by 15,325. Looking at all of this information, we can conclude that the release of the college scorecard did increase the interest in high earning schools based on the large increase in the 'sdindex' variable that we calculated to measure change in terms of search interest. Since it is clear that high earning schools 


